\begin{acknowledgements}
First and foremost I would like to thank my supervisors, Tim French and Rowan
Davies. Tim and Rowan have had a remarkable level of involvement in my honours
project, supporting me greatly in the last year with frequent meetings and
proof-readings of draft papers. It's thanks to them that, in my honours year I
submitted a paper to the Methods for Modalities conference in Osuna, Spain,
which was ultimately accepted for publication, an experience that has further
motivated me to pursue a career in academia. Tim and Rowan must have reviewed
dozens of draft copies of my paper and thesis, providing helpful feedback each
time, and I am truly grateful for the investment of time that they have both
made in supervising my honours project. I would also like to acknowledge the
financial support that Tim and Rowan have granted me, which has allowed me to
attend the conference in Osuna.

I would like to thank John Slaney and Rajeev Gor\'{e} of the Australian National
University, for the Logic Summer School that I attended in 2010. John Slaney
organises the Logic Summer School at ANU each year, and it was thanks to him and
the Australian National University that I received funding to attend the summer
school, where I received my first in-depth introduction to logic. Rajeev
Gor\'{e} in particular gave me my first course in modal logic, from which I
gained a lot, and Rajeev also offered me advice about the topic of my honours
thesis, which I ultimately followed.

I would also like to thank Hans van Ditmarsch, and the anonymous reviewers of
the Methods for Modalities conference that I submitted a paper to this year.
Much of the recommendations that I received from the reviewers, and also from
Hans, has made it into this thesis.

Finally I would like to acknowledge the financial support of the Hackett
Foundation for the Hacket Foundation Alumni Honours Scholarship, which has made
my life easier in the last year, and has partly funded my trip to Osuna for the
Methods for Modalities conference.
\end{acknowledgements}
